\clearpage
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

With the deepening of smart manufacturing and Industry 4.0, constructing high-fidelity and real-time virtual replicas of physical entities has become a core requirement for the implementation of industrial digital twin technologies. In this context, virtual-real mapping serves as a key enabling technology connecting the physical and information worlds, and the quality of its underlying 3D visual reconstruction base directly determines the fidelity and usability of the twin system. Recently, explicit neural rendering technologies represented by 3D Gaussian Splatting (3DGS) have brought a paradigm-shifting breakthrough to novel view synthesis with extremely photorealistic rendering and remarkably high forward rasterization frame rates. However, when directly applied to industrial-like digital twin scenes—which typically contain numerous highly reflective metal components, large areas of textureless solid-color casings, and complex equipment occlusions—traditional 3DGS exposes severe adaptability issues. On the one hand, pure-vision photometric constraints easily fail on complex materials, triggering geometric topological collapse and pervasive "floater" artifacts. On the other hand, the disordered growth of massive discrete Gaussian primitives introduces enormous high-dimensional parameter redundancy and extremely high VRAM consumption, making it difficult to meet the large-scale deployment requirements of edge devices and Web browsers. To address these challenges, this thesis aims to achieve efficient and high-fidelity virtual-real mapping for industrial-like twin scenes by systematically investigating the spatial geometric constraint enhancement and lightweight model compression of 3DGS. The main research work and innovative contributions are as follows:

(1) Proposed a multi-modal adaptive geometric constraint reconstruction algorithm for industrial-like scenes: To address the ill-posed reconstruction problems caused by anisotropic reflections and textureless regions on industrial equipment surfaces, this thesis introduces multi-modal geometric constraints. Initially, through spatial alignment, cross-modal registration is performed between the physical-scale dense point clouds acquired by LiDAR and the relative depth fields inferred by a monocular vision foundation model, injecting a robust 3D physical initial skeleton into 3DGS. Subsequently, an adaptive confidence mask mechanism based on depth local high-frequency feature perception is designed to dynamically partition flat trusted regions and detail-exempted regions. Finally, a loss function combining a Huber depth penalty and first-order normal smoothness regularization is constructed to force the Gaussian primitives to adhere to real physical surfaces under the dynamic scheduling of the mask. Experiments demonstrate that this algorithm effectively suppresses floater noise on metal surfaces and improves geometric tearing on solid-color cabinets. While reducing the 3D geometric error (e.g., Chamfer distance) by approximately 12\%, it well preserves the high-frequency edges of machining textures.

(2) Proposed a lightweight 3DGS representation and edge rendering optimization algorithm for typical twin scenes: Targeting the computing power and VRAM bottlenecks caused by the parameter volume of 3DGS, this thesis tackles the problem from both data reconstruction and rendering pipelines. At the geometric purification level, a multi-dimensional importance assessment model integrating transmittance, global view visibility, and spatial scale is constructed, effectively stripping away redundant "ghost Gaussians" hidden in invisible regions. At the attribute dimension reduction level, Vector Quantization (VQ) technology is introduced to map the high-dimensional Spherical Harmonics (SH) coefficients, which have a large storage overhead, into discrete codebooks. This is combined with a Straight-Through Estimator (STE) to accomplish Quantization-Aware Training (QAT), alleviating the color banding caused by hard quantization. Furthermore, the rendering pipeline adapted for the Web frontend is reconstructed, utilizing WebGPU compute shaders for in-VRAM decoding. Combined with the local sorting optimization of the mobile TBDR architecture, it clears the path for the fluent roaming of lightweight models on ordinary terminals.

(3) Constructed an algorithmic pipeline fusing the improved 3DGS and completed system-level comprehensive verification: This thesis deeply integrates "multi-modal geometric constraints" and "lightweight pruning and quantization" into the training pipeline and designs a staged progressive training strategy. Research shows a good synergistic effect between the two: the stable physical geometric base provides reliable structural guidance for pruning operations. Comprehensive tests on open-source benchmark datasets and a self-collected complex server room twin dataset show that the proposed fused algorithm effectively alleviates the computational and storage pressure of the system while improving image fidelity and geometric accuracy. Compared with traditional methods, the physical storage volume of the scene model is reduced by approximately 12\% (down to about 1010 MB), the peak rendering VRAM is reduced to around 10.2 GB, and the rendering frame rate is increased to 42 FPS under conventional hardware environments. This fused algorithm well balances rendering quality and system overhead, providing a feasible solution for the practical engineering deployment of digital twin systems.

In summary, the research results of this thesis effectively improve the material generalization issues and edge computing power limitations of pure-vision 3DGS in industrial-like scenes. It provides a practical, lightweight, and comprehensive solution for manufacturing enterprises to construct digital twin 3D visual bases with low computing power thresholds, high frame rates, and good fidelity, possessing certain theoretical significance and engineering application prospects.

\vspace{1em}
\noindent\textbf{Keywords:} Industrial Digital Twin; 3D Gaussian Splatting; Multi-modal Geometric Constraints; Lightweight Compression; Virtual-Real Mapping; Adaptive Rendering